---
title: "Statistical Inference Course Project - Part 2"
author: "ACBLimehouse"
date: "7/16/2017"
output:
  pdf_document: default
  html_document:
    keep_md: yes
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=10, fig.height=8, warning=FALSE, message=FALSE)
```

## Overview
The second part of this class project is to perform an analysis of the ToothGrowth data set. The analysis will proceed in four parts:  

1. Load the ToothGrowth data and perform some basic exploratory data analyses 
2. A basic summary of the data.
3. An analysis of variance to compare tooth growth by both supplement and dosage.
4. A statement of conclusions and the necessary assumptions needed for those conclusions. 

## Load Data

```{r}
# load neccesary libraries
library(ggplot2)
library(datasets)
library(gridExtra)
library(GGally)

# The Effect of Vitamin C on Tooth Growth in Guinea Pigs
data(ToothGrowth)
toothGrowth <- ToothGrowth 
toothGrowth$dose <- as.factor(toothGrowth$dose) # convert to factor
```

## Basic Summary of the data
```{r}
str(toothGrowth)
summary(toothGrowth)
head(toothGrowth)

table(toothGrowth$supp, toothGrowth$dose)
```

```{r echo=TRUE}
p1 <- ggplot(data=toothGrowth, aes(x=dose,y=len,fill=dose)) +
  geom_boxplot() + 
  theme(legend.position="none") + 
  facet_grid(.~supp) 

p2 <- ggplot(data=toothGrowth, aes(x=supp,y=len,fill=supp)) +
  geom_boxplot() + 
  theme(legend.position="none") + 
  facet_grid(.~dose) 

p3 <- ggplot(data=toothGrowth, aes(x=supp,y=len,fill=supp)) +
  geom_boxplot()

p4 <- ggplot(data=toothGrowth, aes(x=dose,y=len,fill=dose)) +
  geom_boxplot()

grid.arrange(p1, p4, p2, p3, ncol = 2, nrow=2)
```

## An Analysis of Variance (ANOVA) between supplement and dosage:
```{r}
anova.out <- aov(len ~ supp * dose, data=toothGrowth)
summary(anova.out)
```

The results show there is a notable interaction between the length (len) and dosage (dose) (F(1,54)=15.572;p<0.01)
Also a very clear effect on length(len) by supplement type (supp) (F(2,54)=92;p<0.01). 
Last but not least there is a minor interaction between the combination of supplement type (supp) and dosage (dose) compared to the length (len) (F(2,54)=4.107;p<0.05).

```{r}
TukeyHSD(anova.out)
```

The Tukey HSD  analysis shows that there are significant differences between each of the groups in supp and dose
Only the interactions between VC:0.5-OJ:0.5; VC:1-OJ:0.5; OJ:2-OJ:1; VC:2-OJ:1 and VC:2-OJ:2  are not significant.  

```{r}
confint(anova.out)
print(model.tables(anova.out,"means"),digits=2)
```

## Conclusions

There are clear indications that both the supplement and the dosage have independent effects on the length of guinea pigs' teeth. More, these means indicate longer teeth on avarage. OJ increases avarage growth in combination with dosages 0.5 and 1 more than the VC supplement, while teeth length for the VC supplement vs the OJ in combiantion with dosage 2 has no significant effect (almost same mean & same confidence interval).  

The conclusions are based are based on the following assumptions:

* that the guinea pigs are repesentative for the population of guinea pigs, 
* that dosage and supplement were randomly assigned, and 
* that the distribution of the means is normal.